Relationships in the extremes and their influence on competition and coexistence

### 2021-08-19 paper/Paper.Rnw
draft sent to Dan

methods also provide only a small portion of the avaiable information \citep{Nelsen2006}.

@@

%***DAN: Please combine the following two paragraphs into one single paragraph with this topic sentence:
Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. It is straightforward to see, intuitively, 
why ATAs may influence one mechanism of coexistence, in particular: storage effects. We review, conceptually, what storage effects are and we explain 
why ATAs may be relevant to them.
%***DAN: I have inserted some comments below that might help.
Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
In order for storage effects to contribute as a stabilizing mechanism to coexistence, there must first be covariation between competitive pressures and environmental fluctuations [EC covariance \citep{Chesson_1994, Barabas_2018} ].
%Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
The reliance of storage effect's magnitude on the degree of relationship between environment and competition indicate that storage effects involve environment-coexistence associations, and hence there is potential for ATAs to influence storage effects. 
In addition to EC covariance, functioning storage effects also require that heterospecific populations respond differently to environmental fluctuations, and different life stages must also vary in responses \citep{}. 
In a simple two species system, differences in environmental response allow for one species to grow to high densities in its favorable environemnt while the other is very low. 
Growth is stable for the high density species due to positive EC covariance -- high intraspecific competition limits its growth rate in a favorable environment. 
Conversely, EC covariance is lesser for the low density species, leading to large fluctuations in population growth rate in both directions. 
Although, this may appear like a scenario for extiction, it can, in fact, be advantageous due to differences between life stages, buffering species against extinction by limiting negative growth rates, but not positive ones. 
For example, insensitive adults or dormants seeds can still persist when in an unfavorable environment and overeall growth rate is negative \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.   
Thus, averaged over multiple generations, higher fluctuation in growth rate can lead to a positive low density growth rate and coexistence is possible \citep{Chesson_2000}.
Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships having ATAs, the presence of which may be increasing as climate change brings about more extreme climatic events.

@@

To contextualize idea of how ATAs in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}). The community follows the criteria required for temporal storage effects to operate. The forb dominates and has high fecundity when a sufficient threhold of rain is reached in the early season. The grass dominates and has high fecundity when there is sufficeint rain in the late season. Competition is influenced by resource comsumption of growing juvenile plants. Dormant seeds are unaffected by competiton and unfavorable environment and can persist in the community. When rain fall events temporally fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year. The result is a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated. However, if environmental responses become asymmetrically associated by, say, extreme rainfall events coexistence can be altered. If the community is disturbed by frequent and long floods with some years of normalcy, the grass-forb environmental relationship may become more correlated in the higher values or right-tail associated. Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the lower values or left tail associated. We see the possible consequences of ATA in Figure 2. In simulating their populations,  ATAs can cause species to have longer periods of extreme dominance and lower growth rates overall.

@@

This paper has the following goals. 
G1) To augment the modern theory of coexistence (MTC) to show how ATA plays a role. 
The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
We address how ATA contributes to these components. 
We here examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; spatial variation can be considered in future work.
G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
from ATAs compared to other mechanisms.
G3) To define the ATA rescue or ATA exlusion subset of trait space for which species will coexist or become competitively excluded in our model. 
G4) To provide an empirical example to highlight the role that ATA plays in coexistence and its mechanistic decomposition.
Our examination and separation of the role that ATA play in contributing to coexistence is a novel approach to the study of coexistence...

@@

The other mechanisms in modern coexistence theory that apply in our modelling context of purely temporal variation
make no reference to covariation between quantities, whereas asymmetries of tail associations are 
defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
for which asymmetries of tail association are relevant. Mathematical details supporting this claim are in SI section 
\ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
\cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{i \bs i}$ which is similar to and motivated by the 
original decomposition of \cite{Chesson_1994}, and which we use; the decomposition is how the mechanisms of 
coexistence opperating in a given case are illuminated. The decomposition is 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{i \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{(EC)},\label{eq:previous_decomp}
\end{equation}
where: $\Delta_i^0$ is a baseline;
$\Delta_i^E$ is a contribution to coexistence of environmental variation;
$\Delta_i^C$ is a contribution to coexistence of variation in competition;
$\Delta_i^{(E\shp C)}$ is a contribution to coexistence of having simultaneous variation in both environment and competition, but not including the effects 
of covariation in these quantities; and
$\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- storage effects. 
Only storage effects involve $EC$ covariation,
so the only mechanism of coexistence influenced by ATA is storage effects. Combining our decomposition 
(\ref{eq:storage_effects_further}) with the decomposition (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{i \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[E \nat C]} + \Delta_i^{[EC]},\label{eq:full_decomp}
\end{equation}
which is the extension of the MTC which was goal G1 of the Introduction.

@@

To further explore the model and acheive G3 from the introduction i.e., identifying the ATA rescue or ATA exlusion subset of species trait space for which species will coexist or become competitively excluded in our model, we plotted the invasion growth rate [IGR], $r_{i/i}-q_{12}r_{i/j}$ and the IGR without ATA contributribution  $r_{i/i}-q_{12}r_{i/j} - \Delta_i^{[EC]}$ against the difference between species, $\mu_1-\mu_2$. The same mathematical simplifications described above hold, such that $\mu1$ and $\mu2$ are the only ways that the species differ and computations for coexistence in the lottery model only depend on the difference, $\mu_1-\mu_2$. We henceforward call this difference the trait difference between species or species trait difference or trait space subset range  or species trait range for short. Letting the trait $\mu$ vary within a one dimensional space, this space is then a trait space. Letting there be two points $\mu_1$ and $\mu_2$ within that trait space, then the difference, $\mu_1-\mu_2$, is a range of a subset of that trait space. We then define $\psi_{ATA}$ to be the subset of trait space that allows a postive IGR when ATA is included and $\psi_{~ATA}$ to be the subset of trait space that allows a postive IGR when ATA is not included. We then use symmetric difference $\psi_{ATA} \Delta \psi_{~ATA}$ to make the following conslusions. If the length/range?/rank? of $\psi_{ATA}$ > the length/range?/rank? of $\psi_{~ATA}$ then $\psi_{ATA} \Delta \psi_{~ATA}$ consitutes ATA rescue and the degree of species trait difference that allows ATA rescue to act, hence, the ATA rescue trait space. If the length/range/rank of $\psi_{ATA}$ < the length/range?/rank? of $\psi_{~ATA}$ then $\psi_{ATA} \Delta \psi_{~ATA}$ consitutes ATA exclusion and the degree of species trait difference that allows ATA exclusion to act, hence, the ATA exclusion trait space. In other words, ATA rescue is a subset of all $\mu_1-\mu_2$ that results in coexistence with ATA but failure to coexist without ATA and ATA exclusion is a subset of all $\mu_1-\mu_2$ that results in coexistence without ATA, but failure to coexist with ATA. We highlight and color code these areas in our plots. We made the plots for values of $\delta=0.5, 0.8, 1.0$ and $\sigma=2.0,4.0,5.0, 6.0$. The spaces in between the lines where they intersect the x-axis at y=0, indicate either an ATA rescue trait space or an ATA exclusion trait space. If the line plotting IGR is to the left of the line plotting IGR without ATA, then this space is an ATA rescue trait space. If the line plotting IGR is to the right of the line plotting IGR without ATA, then this space is an ATA exclusion trait space.

@@

\section{Results}\label{section:results}
%from methods : Relative importance of contributions was judged from the plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to the contributions of other mechanisms.
To explore the relative magnitudes of the contributions to coexistence that come from ATA versus other mechanisms in our simple two-species lottery model (G2), we plotted the decomposing mechanisms against increasing $\sigma$ for various values of $\mu_1 - \mu_2$ and $\delta$ (Figure 3). When species are the same, $\mu_1 - \mu_2=0$, only storage effect mechanisms contribute to IGR and other mechanisms (in black) stay at zero. When $\delta=1$, other mechanisms stay relatively more stable with $\Delta_i^0$ set at a negative value and other non-storage effect values being at zero or extremely close to zero. Otherwise, all mechanisms tend to increase in magnitude as $\sigma$ increases. The contribution of ATA captured in $\Delta_i^{[E \nat C]}$ (red) is not small. In almost all cases, aside from when $\delta=1$, ATA contributes negatively to coexistence while symmetric associations, $\Delta_i^{[EC]}$ (blue) contribute postively, and the former decreases with $\sigma$ while the latter increases with $\sigma$. We see the opposite when $\delta=1$, where the contribution of ATA are positve and the contribution of symmetric association are negative and the trends are reversed.

@@

To acheive G3, we assessed the effect of ATA by comparing and plotting (Figure 4) the IGR computed for G2 and the IGR if there were no ATA (IGR - $\Delta_i^{[E \nat C]}$). To find areas of ATA exclusion or ATA rescue we looked at where these line cross y=0 and highlighted and color coded ATA exclusion trait space and ATA rescue trait space (see Methods). In these spaces, contribution of ATA is the mechanism soley responsible for a positive or negative IGR. Occurence of ATA rescue is slim and only occurs when $\delta=1$ and species traits are extremely similar or identical. Other $\delta$ values are more indicative of ATA exclusion events.  When $\delta<1$ ATA exlusion space increase in range? and magnitude as $\sigma$ increases. There is some indication that ATA exlcusion space also increases in magnitude as $\delta$ decreases, however a changes in range have a more complex trend. Range increases across decreasing $\delta$ for all $\sigma$ values except $\sigma=6$. Here, the range is limited by an increasing $\Delta_i^{[E \nat C]}$ as trait space increases and the two IGR's are close to converging.

@@

\begin{figure}
\includegraphics[width=\textwidth]{"../results_figs/fig1.png"}
\caption{
%Figure displaying the noise. Three panels, left-, symmetric, and right-tail assoc noise. Use standard normal marginals.Explain in caption that alternatives for $(b_1,b_2)$ are obtained by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.Jasmin to make and insert the figure here using the includegraphics command, and then draft a caption along the lines described here.
Simple example of bivariate noise with standard normal marginals. A) is left-tail associated, B) is symmetric, and C) is right-tail associated. All three cases have the same Pearson correlation of $\rho=$, up to sampling.., yet show different symmetry of tail associaton. Asymmetrically tail associated cases are perfectly correlated in the lower (A)  or upper (C) halfs of the distribution. The variables $(b_1,b_2)$ are transformed by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.
\label{fig:noise}}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{"../results_figs/figure2.PNG"}
\caption{%Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example. This will be cited from the Intro, so needs to be understandable to someone who does not know any of the ideas we introduce yet. So in particular you cannot use IGR as your metric of coexistence, you need to use something more intuitive. You will pick model parameter values and do two sims, one with ATA noise and one with symmtric noise, and show different outcomes. We should discuss and possibly take several iterations to get this figure right since it's a pedagogically important figure - if the reader is not convinced in the Intro that we are studying something that makes a difference, they are not going to slog through the Theory section, which is inevitably going to be tough going for many readers.
Populations inlfuenced by asymmetric tail association...
recover from low densities less often/ spend more time on average under one-species dominance. A) is a simulation over time of a population within a two species lottery model whose environmental response and competition is influenced by symmetric bivariate noise. B) is like A), but the population is influenced by bivariate noise with left-tailed asymmetric association. Both simulations were ran for total time of 100,000. Noticeable population density was defined at a threshold of 0.95 of total community space. Periods of noticable coexistence are marked in grey. While the population in B) had longer mean periods of coexistence, it had much longer mean periods of one-species dominace. The population in B) also recovered from low densities less often and thus has a lower ability to recover when rare and thus that community has lower coexistence.     
\label{fig:pedagog}}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{"../results_figs/fig3_legend.pdf"}
\caption{The contribution of ATA is non-negligible. All panels show the breakdown of contributing mechanisms of coexistence in accordance to \cite{Ellner_2019} new decomposition method, against increasing $\sigma$. Each panel is defined by a unique set of parameters $\mu_1-\mu_2$ and $\delta$. 
\label{fig:fig3}}
\end{figure}

>>>>

methods also provide only a small portion of the available information \citep{Nelsen2006}.

@@

%***DAN: Please combine the following two paragraphs into one single paragraph with this topic sentence:
Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. It is straightforward to see, intuitively, 
why ATAs may influence one mechanism of coexistence, in particular: storage effects. We review, conceptually, what storage effects are and we explain 
why ATAs may be relevant to them.
%***DAN: I have inserted some comments below that might help.
Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
In order for storage effects to contribute as a stabilizing mechanism to coexistence, there must first be covariation between competitive pressures and environmental fluctuations [EC covariance \citep{Chesson_1994, Barabas_2018} ].
%Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
The reliance of storage effect's magnitude on the degree of relationship between environment and competition indicate that storage effects involve environment-coexistence associations, and hence there is potential for ATAs to influence storage effects. 
In addition to EC covariance, functioning temporal storage effects also require that heterospecific populations respond differently to environmental fluctuations, and different life stages must also vary in responses \citep{}. 
In a simple two species system, differences in environmental response allow for one species to grow to high densities in its favorable environemnt while the other is very low. 
Growth is stable for the high density species due to positive EC covariance -- high intraspecific competition limits its growth rate in a favorable environment. 
Conversely, EC covariance is lesser for the low density species, leading to large fluctuations in population growth rate in both directions. 
Although, this may appear like a scenario for extiction, it can, in fact, be advantageous due to differences between life stages, buffering species against extinction by limiting negative growth rates, but not positive ones. 
For example, insensitive adults or dormants seeds can still persist when in an unfavorable environment and overeall growth rate is negative \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.   
Thus, averaged over multiple generations, higher fluctuation in growth rate can lead to a positive low density growth rate and coexistence is possible \citep{Chesson_2000}.
Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships characterized by ATA, the presence of which may be increasing as climate change brings about more extreme climatic events.

@@

To contextualize idea of how ATA in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}). The community follows the criteria required for temporal storage effects to operate. The forb dominates and has high fecundity when a sufficient threhold of rain is reached in the early season. The grass dominates and has high fecundity when there is sufficeint rain in the late season. Competition is influenced by resource comsumption of growing juvenile plants. Dormant seeds are unaffected by competiton and can persist in the community in an unfavorable environment. When rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year. The result is a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated. However, if environmental responses become asymmetrically associated by, say, extreme rainfall events coexistence can be altered. If the community is disturbed by frequent and long floods with some years of normalcy, the grass-forb environmental relationship may become more correlated in the higher values or right-tail associated. Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the lower values or left tail associated. We see the possible consequences of ATA in Figure 2. In simulating their populations,  ATA can cause species to have longer periods of extreme dominance and lower growth rates overall.

@@

This paper has the following goals. 
G1) To augment the modern theory of coexistence (MTC) to show how ATA plays a role. 
The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
We address how ATA contributes to these components. 
We here examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; spatial variation can be considered in future work.
G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
from ATAs compared to other mechanisms.
G3) To identify and define the ATA rescue or ATA exlusion set of trait space range for which species will coexist or become competitively excluded in our model. 
G4) To provide an empirical example and highlight the role that ATA plays in coexistence and its mechanistic decomposition in a real system.
Our examination and separation of the role that ATA play in contributing to coexistence is a novel approach to the study of coexistence...

@@

The other mechanisms in modern coexistence theory that apply in our modelling context of purely temporal variation
make no reference to covariation between quantities, whereas asymmetries of tail associations are 
defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
for which asymmetries of tail association are relevant. Mathematical details supporting this claim are in SI section 
\ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
\cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$ which is similar to and motivated by the 
original decomposition of \cite{Chesson_1994}, and which we use; the decomposition is how the mechanisms of 
coexistence opperating in a given case are illuminated. The decomposition is 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{(EC)},\label{eq:previous_decomp}
\end{equation}
where: $\Delta_i^0$ is a baseline;
$\Delta_i^E$ is a contribution to coexistence of environmental variation;
$\Delta_i^C$ is a contribution to coexistence of variation in competition;
$\Delta_i^{(E\shp C)}$ is a contribution to coexistence of having simultaneous variation in both environment and competition, but not including the effects 
of covariation in these quantities; and
$\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- storage effects. 
Only storage effects involve $EC$ covariation,
so the only mechanism of coexistence influenced by ATA is storage effects. Combining our decomposition 
(\ref{eq:storage_effects_further}) with the decomposition (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[E \nat C]} + \Delta_i^{[EC]},\label{eq:full_decomp}
\end{equation}
which is the extension of the MTC which was goal G1 of the Introduction.

@@

To further explore the model and acheive G3 from the introduction i.e., to identify and define the ATA rescue or ATA exlusion set of trait space range for which species will coexist or become competitively excluded in our model, we plotted the invasion growth rate [IGR], $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, and the IGR without ATA contributribution, $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}- \Delta_i^{[E \nat C]}$, against the difference between species, $\mu_1-\mu_2$. The same mathematical simplifications described above hold, such that $\mu_1$ and $\mu_2$ are the only ways that the species differ and computations for coexistence in the lottery model only depend on the difference, $\mu_1-\mu_2$. We henceforward call this difference the species trait difference and sets of multiple $\mu_1-\mu_2$ are sets of trait space range. Letting the trait $\mu$ vary within a one dimensional space, this space is then a trait space. Letting there be two points $\mu_1$ and $\mu_2$ within that trait space, then the set of points between $\mu_1$ and $\mu_2$ is a subset of trait space, and the difference, $\mu_1-\mu_2$, is the range of the subset of trait space. We then define $\psi_{ATA}$ to be the subset of trait space bound by a $\mu_1$ and $\mu_2$ whose range defines the maximal species trait difference that allows a postive IGR when ATA is included (coexistence with ATA) and $\psi_{\sim ATA}$ to be the subset of trait space bound by a $\mu_1$ and $\mu_2$ whose range defines the maximal species trait difference that allows a postive IGR when ATA is not included (coexistence without ATA). We then consider the difference between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ to make the following conclusions. 
If the range of $\psi_{ATA}$ > range of $\psi_{\sim ATA}$ then the continuous set of numbers in between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ defines the bounds of extensions to $\psi_{\sim ATA}$ that result coexistence with ATA but failure to coexist without ATA; hence ATA rescue. 
If the range of $\psi_{ATA}$ < the range of $\psi_{\sim ATA}$ then the continuous set of numbers in between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ defines the bounds of extensions to $\psi_{ATA}$ that result coexistence without ATA but failure to coexist with ATA; hence ATA exclusion. 
In other words, ATA rescue is a subset of all $\mu_1-\mu_2$ that results in coexistence with ATA but failure to coexist without ATA, and ATA exclusion is a subset of all $\mu_1-\mu_2$ that results in coexistence without ATA, but failure to coexist with ATA. We highlighted and color coded these areas in our plots. We made the plots for values of $\delta=0.5, 0.8, 1.0$ and $\sigma=2.0,4.0,5.0, 6.0$. The spaces in between the lines where they intersect the x-axis at y=0, indicate either an ATA rescue trait space or an ATA exclusion trait space. If the line plotting IGR is to the left of the line plotting IGR without ATA, then this space is an ATA rescue trait space. If the line plotting IGR is to the right of the line plotting IGR without ATA, then this space is an ATA exclusion trait space.

@@

\section{Results}\label{section:results}
%from methods : Relative importance of contributions was judged from the plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to the contributions of other mechanisms.
To explore the relative magnitudes of the contributions to coexistence that come from ATA versus other mechanisms in our simple two-species lottery model (G2), we plotted the decomposing mechanisms against increasing $\sigma$ for various values of $\mu_1 - \mu_2$ and $\delta$ (Figure 3). When species are the same, $\mu_1 - \mu_2=0$, only storage effect mechanisms contribute to IGR and other mechanisms (in black) stay at zero. When $\delta=1$, other mechanisms stay relatively more stable with $\Delta_i^0$ set at a negative value and other non-storage effect values being at zero or extremely close to zero. Otherwise, all mechanisms tend to increase in magnitude as $\sigma$ increases. The contribution of ATA captured by $\Delta_i^{[E \nat C]}$ (red) is not small. In almost all cases, aside from when $\delta=1$, ATA contributes negatively to coexistence while symmetric associations, $\Delta_i^{[EC]}$ (blue) contribute postively, and the former decreases with $\sigma$ while the latter increases with $\sigma$. We see the opposite when $\delta=1$, where the contribution of ATA are positve and the contribution of symmetric association are negative and the trends are reversed.

@@

To acheive G3, we assessed the effect of ATA by comparing and plotting (Figure 4) the IGR computed for G2 and the IGR if there were no ATA (IGR - $\Delta_i^{[E \nat C]}$). To find areas of ATA exclusion or ATA rescue we looked at where these line cross y=0 and highlighted and color coded ATA exclusion trait space and ATA rescue trait space (see Methods). In these spaces, contribution of ATA is the mechanism soley responsible for a positive or negative IGR. Occurence of ATA rescue is slim and only occurs when $\delta=1$ and species traits are extremely similar or identical. Other $\delta$ values are more indicative of ATA exclusion events.  When $\delta<1$ ATA exlusion space increase in range and magnitude as $\sigma$ increases. There is indication that ATA exlcusion space also increases in magnitude as $\delta$ decreases, however changes in range have a more complex trend. Range increases across decreasing $\delta$ for all $\sigma$ values except $\sigma=6$. Here, the range is limited by an increasing $\Delta_i^{[E \nat C]}$ as trait space increases and the two IGR's are close to converging.

@@

\begin{figure}
\includegraphics[width=\textwidth]{"../results_figs/fig1.png"}
\caption{
%Figure displaying the noise. Three panels, left-, symmetric, and right-tail assoc noise. Use standard normal marginals.Explain in caption that alternatives for $(b_1,b_2)$ are obtained by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.Jasmin to make and insert the figure here using the includegraphics command, and then draft a caption along the lines described here.
Simple example of bivariate noise with standard normal marginals. A) is left-tail associated, B) is symmetric, and C) is right-tail associated. All three cases have the same Pearson correlation up to sampling variation, yet show different symmetry of tail associaton. Asymmetrically tail associated cases are perfectly correlated in the lower (A)  or upper (C) halfs of the distribution. The variables $(b_1,b_2)$ are transformed by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.
\label{fig:noise}}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{"../results_figs/figure2.PNG"}
\caption{%Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example. This will be cited from the Intro, so needs to be understandable to someone who does not know any of the ideas we introduce yet. So in particular you cannot use IGR as your metric of coexistence, you need to use something more intuitive. You will pick model parameter values and do two sims, one with ATA noise and one with symmtric noise, and show different outcomes. We should discuss and possibly take several iterations to get this figure right since it's a pedagogically important figure - if the reader is not convinced in the Intro that we are studying something that makes a difference, they are not going to slog through the Theory section, which is inevitably going to be tough going for many readers.
Populations inlfuenced by asymmetric tail association recover from low densities less often/ spend more time on average under one-species dominance. The top panel, A, is a simulation over time of a population within a two species lottery model whose environmental response and competition is influenced by symmetric bivariate noise. The bottom panel, B, is like A, but the population is influenced by bivariate noise with left-tailed asymmetric association. Both simulations were ran for total time of 100000. Noticeable population density was defined at a threshold of 0.95 of total community space. Periods of noticable coexistence are marked in grey. While the population in B had longer mean periods of coexistence, it had much longer mean periods of one-species dominace. The population in B also recovered from low densities less often and thus has a lower ability to recover when rare and thus that community has lower coexistence.     
\label{fig:pedagog}}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{"../results_figs/fig3_legend.pdf"}
\caption{The contribution of ATA is non-negligible. All panels show the breakdown of contributing mechanisms of coexistence in accordance to \cite{Ellner_2019} new decomposition method, against increasing $\sigma$. Each panel is defined by a unique set of parameters $\mu_1-\mu_2$ and $\delta$. In most cases, the contribution of ATA (red) is negatively; however when $\delta=1$, ATA contribution is positive. The contribution of ATA is non-negligible and in special cases result in a sign change of IGR (orange). 
\label{fig:fig3}}
\end{figure}
